\documentclass[12pt, a4paper]{article}

\usepackage{lipsum}

\usepackage{indentfirst}
\usepackage{parskip}
\usepackage{setspace}
\setlength{\parskip}{0px}
\setlength{\parindent}{1.25cm}
\doublespacing


\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhf{}
\fancyhead[R]{\thepage}



\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{blindtext}


\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage{wrapfig}
\usepackage{float}


\usepackage{multirow}
\usepackage{array}
\usepackage{tabu}
\titleformat{\section}
{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
{\normalfont\large\bfseries}{\thesubsection}{1em}{}


\counterwithin{equation}{section}

\usepackage{hyperref}
\urlstyle{same}



\title{ IB Mathematics AA Higher Level\\
    
    Exploring and implementing 3D graphics through linear
algebra }
\author{Lars Sawadsky}
\date{\today}
\begin{document}

    \begin{titlepage}
        \begin{center}
            \vspace*{9cm}

            IB Mathematics AA Higher Level\\
            \medskip
            Exploring and implementing 3D computer graphics
            using linear algebra

            \vspace{8cm}

            L. Sawadsky\\
            Internal Assessment: Mathematics AA HL\\
            Session: May 2026\\
            Word Count:
        \end{center}
        \newpage
    \end{titlepage}

    \tableofcontents
    \newpage

    \section{Introduction}

    Representing our three-dimensional world in the two
    dimensions our computer screens allow us has proven not to
    be a simple task, but thankfully one that centuries of
    research in linear algebra happen to assist greatly in.
    Simulating a three-dimensional environment in a computer
    system is done largely using research in physics: laws and
    constants that govern the behaviour of rigid bodies, fluid,
    and light are replicated in programs to create a world that
    functions, more or less, like ours. However, there is a
    problem designers run into that has no parallel in our
    world. We understand the laws by which physical things
    change, but \textit{representing} these changes is a
    challenge unique to computerized versions of the world. My
    interests in video games, many of which heavily rely on 3D
    graphics algorithms, and computer science have led me to
    want to explore the mathematics and computation behind
    graphical engines. To support this investigation, I will
    also develop an implementation of the described principles
    in C++, using the 2D graphics library SDM to create a simple
    3D graphics library. This investigation will explore two aspects
    of 3D graphics: the projection of a set of given
    vertices in 3D space onto a 2D screen, and some common
    manipulations of those vertices: translations, stretches,
    and rotations.

    \subsection{Context}

    Modern 3D graphics engines, such as OpenGL, process graphics
    in a pipeline: every refresh, operations are taken on space
    and the camera and the display is updated. Each
    transformation and camera update is processed as a 4x4
    transformation matrix (this will be expanded on later in the
    investigation), and they are then multiplied together into
    one 

    \section{}

\end{document}